A cascaded supervised learning approach to inverse reinforcement learning

### 2013-03-27 piot2012cascading/ECML13_cascading.tex
Version ECML envoyée à M&O

\begin{abstract}
  This paper considers the Inverse Reinforcement Learning (IRL) problem, that is inferring a reward function for which a demonstrated expert policy is optimal.
We propose to break the IRL problem down into two generic Supervised Learning steps: this is the Cascaded Supervised IRL (CSI) approach. A classification step that defines a score function is followed by a regression step providing a reward function.
A theoretical analysis shows that the demonstrated expert policy is near-optimal for the computed reward function.
Not needing to repeatedly solve a Markov Decision Process (MDP) and the ability to leverage existing techniques for classification and regression are two important advantages of the CSI approach. It is furthermore empirically demonstrated to work with only transitions sampled according to the expert policy, up to the use of some heuristics. This is exemplified on two classical benchmarks (a highway driving simulator and an inverted pendulum simulator).
  \end{abstract}
\section{Introduction}
\label{sec-2}
Sequential decision making consists in choosing the appropriate action given the available data in order to maximize a certain criterion. When framed in a Markov Decision Process (MDP) (see Sec.~\ref{sec:background}), (Approximate) Dynamic programming ((A)DP) or Reinforcement Learning (RL) are often used to solve the problem by maximizing the expected sum of discounted rewards. The Inverse Reinforcement Learning (IRL)~\cite{russell1998learning} problem, which is addressed here, aims at inferring a reward function for which a demonstrated expert policy is optimal.

@@

This algorithm does not need to iteratively solve an MDP and needs only sampled transitions from expert and non-expert policies as inputs. Moreover, up to the use of some heuristics (see Sec.~\ref{subsubsec:heuristics}), the algorithm is able to be trained only with transitions sampled from the demonstrated expert policy. A specific instantiation of CSI (proposed Sec.~\ref{subsec:instanciation}) is tested on an inverted pendulum simulator (Sec.~\ref{subsec:IP}) and a car driving simulator (Sec.~\ref{subsec:highway}) where we compare it with an AL algorithm \cite{abbeel2004apprenticeship}, with a pure classification algorithm \cite{taskar2005learning} and with a recent successfull IRL method \cite{klein2012scirl} as well as with a random baseline.

@@

We work with finite MDPs \cite{puterman1994markov}, that is tuples $\{S,A,P,R,\gamma\}$. The state space is noted $S$, $A$ is a finite action space, $R\in\mathbb{R}^{S\times A}$ is a reward-function, $\gamma\in (0,1)$ is a discount factor and $P\in \Delta_{S}^{S\times A}$ is the Markovian dynamics of the MDP. Thus, for each $(s,a)\in S\times A$, $P(.|s,a)$ is a distribution over $S$ and $P(s'|s,a)$ is the probability to reach $s'$ by choosing action $a$ in state $s$. At each time step $t$, the agent uses the information encoded in the state $s_t\in S$ in order to choose an action $a_t \in A$ according to a (deterministic\footnote{We restrict ourselves here to deterministic policies, but the loss of generality is minimal as there exists at least one optimal deterministic policy.}) policy $\pi\in A^S$. The agent then steps to a new state $s_{t+1}\in S$ according to the Markovian transition probabilities $P(s_{t+1}|s_t,a_t)$. Given that $P_\pi$ is $(P(s'|s,\pi(s)))_{s,s' \in S}$, the irreductible transition probability matrix, the stationary distribution over the states $\rho_\pi$ induced by a policy $\pi$ satisfies
\begin{equation}
  \rho_\pi^TP_\pi = \rho_\pi^T\textrm{, with }X^T\textrm{ being the transpose of }X.
  \end{equation}
The stationary distribution relative to the exert policy $\pi_E$ is $\rho_E$.

@@

We recall that solving the MDP is the direct problem. This contribution aims at solving the inverse one. We observe trajectories drawn from an expert's deterministic\footnotemark[\value{footnote}] policy $\pi_E$, assuming that there exists some unknown reward $R_E$ for which the expert is optimal. The suboptimality of the expert is an interesting setting that has been discussed for example in \cite{melo2010analysis,syed2010reduction}, but that we are not addressing here. We do not try to find this unknown reward $R_E$ but rather a non trivial reward $R$ for which the expert is at least near-optimal. The trivial reward $0$ is a solution to this ill-posed problem (no reward means that every behavior is optimal). Because of its ill-posed nature, this expression of \emph{Inverse Reinforcement Learning} (IRL) still has to find a satisfactory solution although a lot of progress have been made, see Sec.~\ref{sec:related}.
\section{The cascading algorithm}
\label{sec:algo}
Our first step towards a reward function solving the IRL problem is a classification step using a \emph{score function-based multi-class classifier} (SFMC$^2$ for short). This classifier learns a score function $q\in\mathbb{R}^{S\times A}$ that rates the association of a given action\footnote{Here, actions play the role of what is known as {\it labels} or {\it categories} when talking about classifiers.} $a\in A$ with a certain input $s\in S$. The classification rule $\pi_C\in A^S$ simply selects (one of) the action(s) that achieves the highest score for the given inputs:
\begin{equation}
  \label{eq:greedy2}
\pi_C(s) \in \argmax_a q(s,a).
\end{equation}
For example, \emph{Multi-class Support Vector Machines}~\cite{guermeur2011generic} can be seen as SFMC$^2$ algorithms, the same can be said of the structured margin approach \cite{taskar2005learning} which we consider in the experimental setting. Other algorithms may be envisionned (see Sec.~\ref{subsec:instanciation}).

Given a dataset $D_C = \{(s_i,a_i=\pi_E(s_i))_i\}$ of actions $a_i$ (deterministically) chosen by the expert on states $s_i$, we train such a classifier. The classification policy $\pi_C$ is not the end product we are looking for (that would be mere imitation of the expert, not IRL). What is of particular interest to us is the score function $q$ itself. One can easily notice the similarity between Eq.~\eqref{eq:greedy2} and Eq.~\eqref{eq:greedy} that describes the relation between the optimal policy in an MDP and its optimal action-value function. The score function $q$ of the classifier can thus be viewed as some kind of optimal action-value function for the classifier policy $\pi_C$. By inversing the Bellman equation Eq.~\eqref{eq:bellman1} with $q$ in lieu of $Q^\pi_R$, one gets $R^C$, the reward function relative to our score/action-value function $q$:
\begin{equation}
  \label{eq:rc}
  R^C(s,a) =q(s,a) - \sum_{\mathrm{s'}}P(s'|s,a)q(s',\pi_C(s')).
\end{equation}
As we wish to solve the general, or approximate, IRL problem where the transition probabilities $P$ are unknown, our reward function $R^C$ will be approximated with the help of information gathered by interacting with the system. We assume that another dataset $D_R = \{(s_j,a_j,s'_j)_j\}$ is available where $s'_j$ is the state an agent taking action $a_j$ in state $s_j$ transitioned to. Action $a_j$ need not be chosen by any particular policy. The dataset $D_R$ brings us information about the dynamics of the system. From it, we construct datapoints
\begin{equation}
  \label{eq:rj}
  \{\hat r_j = q(s_j,a_j) - \gamma q(s'_j,\pi_C(s'_j))\}_j.
\end{equation}
As $s'_j$ is sampled accordingly to $P(\cdot|s_j,a_j)$ the constructed datapoints help building a good approximation of $R^C(s=s_j,a=a_j)$. A regressor (a simple least-square approximator can do but other solutions such as Gaussian Processes \cite{rasmussen2006gaussian} could also be envisionned) is then fed the datapoints $((s_j,a_i),\hat r_j)$ to obtain $\hat R^C$, a generalisation of $\{((s_j,a_j),\hat r_j)_j\}$ over the whole state-action space. The complete algorithm is given in Alg.~\ref{algo:cascading}.

@@

Readers familiar with the work presented in \cite{klein2012scirl} will see some similarities between the theoretical analyses of SCIRL and CSI. As both algorithms solve the same problem, both theorems naturally bound the same criterion. Another shared feature is the use of the score function $q$ of the classifier as a proxy for the action-value function of the expert. The same structure can therefore be found in both analyses as there is only so many ways to study error propagation in IRL algorithms.\\

@@

First let's introduce some notations, $q\in\mathbb{R}^{S\times A}$ is the score function output by the classification step, $\pi_C$ is a deterministic classifier policy so that $\forall s \in S,\pi_C(s)\in\argmax_{a\in A}q(s,a)$. We recall the definition of the reward $R^C\in\mathbb{R}^{S\times A}$:

@@

and the definition of $\hat{R}^C\in\mathbb{R}^{S\times A}$ which is the reward function output by the regression step.

The difference between $R^C$ and $\hat{R}^C$ is noted $\epsilon^R=R^C-\hat{R}^C$.
We also introduce the reward function $R^E\in\mathbb{R}^{S\times A}$ which will be useful in our proof:
\begin{equation}
\forall (s,a)\in S\times A, R^E(s,a)=q(s,a) -\gamma\sum_{s'\in S}P(s'|s,a)q(s',\pi_E(s')).
\end{equation}

@@

Other differences between existing IRL or apprenticeship learning approaches and the proposed cascading algorithm are further examined in the following section.

%% Feel good comment
\section{Experiments}
\label{sec:experiments}
In this section, we put our approach under stress to empirically demonstrate its behavior. We begin by providing information pertaining to both benchmarks: an explanation about the amount and source of the available data, the rationale behing the heuristics we use to compensate for the dire data scarcity and a quick word about the contender CSI is compared to are given Sec.~\ref{subsec:generalities}. We supply quantitative results and comparison of CSI with state-of-the art approaches on first a classical RL benchmark (the mountain car) %FIXME: subsec moutnain car
and then on a more difficult highway driving simulator (Sec.~\ref{subsec:highway}).
\subsection{Generalities}
\label{subsec:generalities}
\subsubsection{Data scarcity}
\label{subsubsec:data}
The CSI algorithm was designed not to repeatedly solve the RL problem. This feature makes it particularly well-suited to environments where sampling the MDP is difficult or costly. In the experiments, CSIis fed only with data sampled according to the expert policy. This corresponds for example to a situation where a costly system can only be controlled by a trained operator as a bad control sequence could lead to a system breakdown.

More precisely, the expert controls the system for $M$ runs of lengths ${L_i}_{1\leq i \leq M}$, %with values of $n$ in $\{3,7,10,15,20\}$,
giving samples $(s_k,a_k=\pi_E(a_k),s'_k)_k = D_E$. The dataset $D_C$ fed to the classifier is straightforwardly constructed from $D_E$ by dropping the $s'_k$ terms:
\begin{equation}
  D_C = \{(s_i = s_k,a_i=a_k)_i\}.
\end{equation}
\subsubsection{Heuristics}
\label{subsubsec:heuristics}

There is no particuliar problem for the classification step. It is however not reasonable to construct the dataset $D_R = \{((s_k,a_k=\pi_E(s_k)),\hat r_k)_k\}$ only from expert transitions and expect a small regression error term $\epsilon_R$. Indeed, the dataset $D_E$ only samples the dynamics induced by the expert's policy and not the whole dynamics of the MDP. 
This means that for a certain state $s_k$ we only know the corresponding expert action $a_k=\pi_E(s_k)$ and the following state $s'_k$ sampled according to the MDP dynamics : $s'_k \sim P(\cdot|s_k,a_k)$. For the regression to be meaningful, we need samples associating the same state $s_k$ and a different action $a\neq a_k$ with a datapoint $\hat r \neq \hat r_k$.\\
Recall that $\{\hat r_j = q(s_j,a_j) - \gamma q(s'_j,\pi_C(s'_j))\}_j$ (Eq.~\ref{eq:rj}), without knowing $s'_k \sim P(\cdot|s_k,a\neq a_k)$, we can not provide the regressor with a datapoint to associate with $(s_k,a\neq a_k)$.
We will artificially augment the dataset $D_R$ with samples $((s_j=s_k,a),r_{min})_{j;\forall a\neq \pi_E(s_j) = a_k}$ where
\begin{equation}
  r_{min} = \min_j\hat r_j - 1.
\end{equation}
This heuristics intructs the regressor to associate a state-action tuple disagreeing with the expert ({\it i.e. }$(s_k,a\neq a_k)$) with a reward strictly inferior to any of those associated with expert state action tuples ({\it i.e. }$(s_k,a_k=\pi_E(s_k))$). Semantically, we are asserting that disagreeing with the expert in states the expert visited is a bad idea, thus the shape of the reward function should reflect it. This heuristics says nothing about states absent from the expert dataset. For such states the generalization capabilities of the regressor and, later on, the exploration of the MDP by an agent optimizing the reward will solve the problem.
Although this heuristics was not analyzed in Sec.~\ref{sec:analysis} (where the availability of a more complete dataset $D_R$ was assumed), the results shown in the next two subsections demonstrate its soundness.
\subsubsection{Comparison with state of the art approaches}
The similar looking yet fundamentally different algorithm SCIRL is an obvious choice as a contender to CSI as it advertises the same ability to work with very little data, without repeatedly solving the RL problem. In both experiments we give the exact same data to CSI and SCIRL.

@@

Finally, the classification policy $\pi_C$ output by the classification step of CSI was evaluated as well. Comparing classification and IRL algorithms makes no sense if the object of interest is the reward itself as can be envisionned in a biological or economical context. It is however perfectly sound to do so in an imitation context where what matters is the performance of the agent with respect to some objective criterion. Both experiments use such a criterion. Calssification algorithms don't have to optimize any reward, the classification policy can directly be used in the environment. IRL algorithms, on the other hand, output a reward that must then be plugged in a RL or DP algorithm to get a policy. In each benchmark we used the same (benchmark-dependant) algorithm to get a policy from each of the three rewards output by SCIRL, CSI and RE. It is these policies whose performance we show. Finding a policy from a reward if of course a non-trivial problem that should not be swept under the rug, nevertheless we choose not to concern ourselves with it here as we wish to focus on IRL algorithms, not RL or DP algorithms. In this regard, using a classifier that directly outputs a policy may seem a much simpler solution, but we hope that the reader will be convinced that the gap in performance between classification and IRL is worth the trouble.

@@

We referred in Sec.\ref{sec:algo} to \emph{score-function based multi-class classifiers} and explained how the classification rule is similar to the greedy mechanism that exists between an optimal action-value function and an optimal policy in an MDP. Most classification algorithm can be seen as such a classifier. In a simple $k$-nearest neighboor approach, for example, the score function $q(s,a)$ is the number of elements of class $a$ among the $k$-nearest neighboors of $s$. The generic M-SVM model makes the score function explicit (see \cite{guermeur2011generic}) (we use a SVM in the mountain car experiment Sec.~\ref{sec:mountaincar}). In the highway experiment, we choose to use a structured margin classification approach \cite{taskar2005learning} that, given a feature vector $\phi: S\times A \rightarrow \mathbb{R}^p$ solves:
\begin{eqnarray*}
  &\min_{\theta,\zeta}\frac{1}{2}\|\theta\|^2 +
  \frac{\eta}{N}\sum_{i=1}^N \zeta_i \\
  &\text{~s.t.~} \forall i,
  \theta^T{\phi}(s_i,a_i)+\zeta_i \geq \max_a \theta^T
  {\phi}(s_i,a) + l(s_i,a). \label{eq:qp_taskar}
\end{eqnarray*}
with $l(s_i,a) = 1\textrm{ if } a\neq a_i \textrm{ and } l(s_i,a_i)=0$. The score function of this classifier is simply $q(s,a) = \theta^T \phi(s,a)$. We chose a SVR as a regressor in the mountain car experiment and a simple least-square regressor on the highway. Using the same features $\phi$ as our classifier, we introduce $\lambda$ a regularization coefficient (set to $0.1$ in our experiments), $I$ the identity matrix and 
\begin{eqnarray}
  \Phi = \begin{bmatrix}
    \vdots\\
    \phi(s_j,a_j)\\
    \vdots
  \end{bmatrix} \in \mathbb{R}^{n|A|\times p}\textrm{ and }
  \hat R = \begin{bmatrix}
    \vdots\\
    \hat r_j\\
    \vdots
  \end{bmatrix}\in \mathbb{R}^{n|A|},
\end{eqnarray}
with $r_j = r_{min}$ if $a_j \neq \pi_E(s_j)$. Our regressor can be written:
\begin{equation}
  \omega = (\Phi^T\Phi + \lambda I)^{-1}\Phi^T\hat R.
\end{equation}
The reward function is: $\hat R^C(s,a) = \omega^T \phi(s,a)$

@@

The feature function $\phi$ for this setting is the natural tabular feature function where each state action couple $(s,a)$ is associated with a unique index $i_{(s,a)}$. The only non-zero component of $\phi(s,a)$ is then the component at index $i_{(s,a)}$.

@@

Expert performance averages to $7.74$ ; we also show the natural random baseline consists in drawing a random reward vector (with a uniform law) and training an agent on it. The reward functions found by SCIRL, CSI and RE are then optimized using a DP algorithm. The dataset $D_{random}$ needed by RE (and neither by CSI nor SCIRL) is made of $100$ episodes of length $10$ starting randomly in the state space and following a random policy. the dataset $D_E$ is made of $n$ episodes of length $n$, with $n \in \{3,7,10,15,20\}$.

Results are shown Fig.~\ref{fig:Highway} and Tab.~\ref{tab:highway}. We give the values of $\mathbf{E}_{s\sim\mathcal{U}}[V^{\pi}_{R_E}(s)]$ with $\pi$ being in turn the optimal policy for the rewards given by SCIRL, CSI and RE, the policy $\pi_C$ of the classifier (the very one the classification step of CSI outputs), the expert's policy $\pi_E$ and the optimal policy for a randomly drawn reward. Performance for CSI is slightly but definitively higher than for SCIRL (see the variance in Tab.~\ref{tab:highway}), slightly below the performance of the expert itself. Very few samples (100) are needed to reliably achieve expert-level performance.

>>>>

\begin{abstract}
  This paper considers the Inverse Reinforcement Learning (IRL) problem, that is inferring a reward function for which a demonstrated expert policy is optimal.
We propose to break the IRL problem down into two generic Supervised Learning steps: this is the Cascaded Supervised IRL (CSI) approach. A classification step that defines a score function is followed by a regression step providing a reward function.
A theoretical analysis shows that the demonstrated expert policy is near-optimal for the computed reward function.
Not needing to repeatedly solve a Markov Decision Process (MDP) and the ability to leverage existing techniques for classification and regression are two important advantages of the CSI approach. It is furthermore empirically demonstrated to work with only transitions sampled according to the expert policy, up to the use of some heuristics. This is exemplified on two classical benchmarks (the mountain car problem and a highway driving simulator).
  \end{abstract}
\section{Introduction}
\label{sec-2}
Sequential decision making consists in choosing the appropriate action given the available data in order to maximize a certain criterion. When framed in a Markov Decision Process (MDP) (see Sec.~\ref{sec:background}), (Approximate) Dynamic programming ((A)DP) or Reinforcement Learning (RL) are often used to solve the problem by maximizing the expected sum of discounted rewards. The Inverse Reinforcement Learning (IRL)~\cite{russell1998learning} problem, which is addressed here, aims at inferring a reward function for which a demonstrated expert policy is optimal.

@@

This algorithm does not need to iteratively solve an MDP and needs only sampled transitions from expert and non-expert policies as inputs. Moreover, up to the use of some heuristics (see Sec.~\ref{subsubsec:heuristics}), the algorithm is able to be trained only with transitions sampled from the demonstrated expert policy. A specific instantiation of CSI (proposed Sec.~\ref{subsubsec:subroutines}) is tested on the mountain car problem (Sec.~\ref{subsec:mountaincar}) and a highway driving simulator (Sec.~\ref{subsec:highway}) where we compare it with a pure classification algorithm \cite{taskar2005learning} and with two recent successfull IRL methods \cite{klein2012scirl} as well as with a random baseline.

@@

We work with finite MDPs \cite{puterman1994markov}, that is tuples $\{S,A,P,R,\gamma\}$. The state space is noted $S$, $A$ is a finite action space, $R\in\mathbb{R}^{S\times A}$ is a reward-function, $\gamma\in (0,1)$ is a discount factor and $P\in \Delta_{S}^{S\times A}$ is the Markovian dynamics of the MDP. Thus, for each $(s,a)\in S\times A$, $P(.|s,a)$ is a distribution over $S$ and $P(s'|s,a)$ is the probability to reach $s'$ by choosing action $a$ in state $s$. At each time step $t$, the agent uses the information encoded in the state $s_t\in S$ in order to choose an action $a_t \in A$ according to a (deterministic\footnote{We restrict ourselves here to deterministic policies, but the loss of generality is minimal as there exists at least one optimal deterministic policy.}) policy $\pi\in A^S$. The agent then steps to a new state $s_{t+1}\in S$ according to the Markovian transition probabilities $P(s_{t+1}|s_t,a_t)$. Given that $P_\pi$ is $(P(s'|s,\pi(s)))_{s,s' \in S}$ the irreductible transition probability matrix, the stationary distribution over the states $\rho_\pi$ induced by a policy $\pi$ satisfies
\begin{equation}
  \rho_\pi^TP_\pi = \rho_\pi^T\textrm{, with }X^T\textrm{ being the transpose of }X.
  \end{equation}
The stationary distribution relative to the exert policy $\pi_E$ is $\rho_E$.

@@

We recall that solving the MDP is the direct problem. This contribution aims at solving the inverse one. We observe trajectories drawn from an expert's deterministic\footnotemark[\value{footnote}] policy $\pi_E$, assuming that there exists some unknown reward $R_E$ for which the expert is optimal. The suboptimality of the expert is an interesting setting that has been discussed for example in \cite{melo2010analysis,syed2010reduction}, but that we are not addressing here. We do not try to find this unknown reward $R_E$ but rather a non trivial reward $R$ for which the expert is at least near-optimal. The trivial reward $0$ is a solution to this ill-posed problem (no reward means that every behavior is optimal). Because of its ill-posed nature, this expression of \emph{Inverse Reinforcement Learning} (IRL) still has to find a satisfactory solution although a lot of progress have been made, see Sec.~\ref{sec:related}.
\section{The cascading algorithm}
\label{sec:algo}
Our first step towards a reward function solving the IRL problem is a classification step using a \emph{score function-based multi-class classifier} (SFMC$^2$ for short). This classifier learns a score function $q\in\mathbb{R}^{S\times A}$ that rates the association of a given action\footnote{Here, actions play the role of what is known as {\it labels} or {\it categories} when talking about classifiers.} $a\in A$ with a certain input $s\in S$. The classification rule $\pi_C\in A^S$ simply selects (one of) the action(s) that achieves the highest score for the given inputs:
\begin{equation}
  \label{eq:greedy2}
\pi_C(s) \in \argmax_a q(s,a).
\end{equation}
For example, \emph{Multi-class Support Vector Machines}~\cite{guermeur2011generic} can be seen as SFMC$^2$ algorithms, the same can be said of the structured margin approach \cite{taskar2005learning} both of which we consider in the experimental setting. Other algorithms may be envisionned (see Sec.~\ref{subsubsec:subroutines}).

Given a dataset $D_C = \{(s_i,a_i=\pi_E(s_i))_i\}$ of actions $a_i$ (deterministically) chosen by the expert on states $s_i$, we train such a classifier. The classification policy $\pi_C$ is not the end product we are looking for (that would be mere imitation of the expert, not IRL). What is of particular interest to us is the score function $q$ itself. One can easily notice the similarity between Eq.~\eqref{eq:greedy2} and Eq.~\eqref{eq:greedy} that describes the relation between the optimal policy in an MDP and its optimal action-value function. The score function $q$ of the classifier can thus be viewed as some kind of optimal action-value function for the classifier policy $\pi_C$. By inversing the Bellman equation Eq.~\eqref{eq:bellman1} with $q$ in lieu of $Q^\pi_R$, one gets $R^C$, the reward function relative to our score/action-value function $q$:
\begin{equation}
  \label{eq:rc}
  R^C(s,a) =q(s,a) - \sum_{\mathrm{s'}}P(s'|s,a)q(s',\pi_C(s')).
\end{equation}
As we wish to solve the general, or approximate, IRL problem where the transition probabilities $P$ are unknown, our reward function $R^C$ will be approximated with the help of information gathered by interacting with the system. We assume that another dataset $D_R = \{(s_j,a_j,s'_j)_j\}$ is available where $s'_j$ is the state an agent taking action $a_j$ in state $s_j$ transitioned to. Action $a_j$ need not be chosen by any particular policy. The dataset $D_R$ brings us information about the dynamics of the system. From it, we construct datapoints
\begin{equation}
  \label{eq:rj}
  \{\hat r_j = q(s_j,a_j) - \gamma q(s'_j,\pi_C(s'_j))\}_j.
\end{equation}
As $s'_j$ is sampled accordingly to $P(\cdot|s_j,a_j)$ the constructed datapoints help building a good approximation of $R^C(s=s_j,a=a_j)$. A regressor (a simple least-square approximator can do but other solutions could also be envisionned, see Sec.~\ref{subsubsec:subroutines}) is then fed the datapoints $((s_j,a_i),\hat r_j)$ to obtain $\hat R^C$, a generalisation of $\{((s_j,a_j),\hat r_j)_j\}$ over the whole state-action space. The complete algorithm is given in Alg.~\ref{algo:cascading}.

@@

Readers familiar with the work presented in \cite{klein2012scirl} will see some similarities between the theoretical analyses of SCIRL and CSI. As both algorithms solve the same problem, both theorems naturally bound the same criterion. Another shared feature is the use of the score function $q$ of the classifier as a proxy for the action-value function of the expert $Q^{\pi_E}_{R_E}$. The same structure can therefore be found in both analyses as there is only so many ways to study error propagation in IRL algorithms.\\

@@

First let's recall some notations, $q\in\mathbb{R}^{S\times A}$ is the score function output by the classification step, $\pi_C$ is a deterministic classifier policy so that $\forall s \in S,\pi_C(s)\in\argmax_{a\in A}q(s,a)$, $R^C\in\mathbb{R}^{S\times A}$ is:

@@

and $\hat{R}^C\in\mathbb{R}^{S\times A}$ is the reward function output by the regression step.

The difference between $R^C$ and $\hat{R}^C$ is noted $\epsilon^R=R^C-\hat{R}^C$.
We also introduce the reward function $R^E\in\mathbb{R}^{S\times A}$ which will be useful in our proof, not to be confused with $R_E$ the unknown reward function the expert optimizes:
\begin{equation}
\forall (s,a)\in S\times A, R^E(s,a)=q(s,a) -\gamma\sum_{s'\in S}P(s'|s,a)q(s',\pi_E(s')).
\end{equation}

@@

Other differences between existing IRL or apprenticeship learning approaches and the proposed cascading algorithm are further examined Sec.~\ref{sec:related}.

%% Feel good comment
\section{Experiments}
\label{sec:experiments}
In this section, we put our approach under stress to empirically demonstrate its behavior. We begin by providing information pertaining to both benchmarks: an explanation about the amount and source of the available data, the rationale behing the heuristics we use to compensate for the dire data scarcity and a quick word about the contender CSI is compared to are given Sec.~\ref{subsec:generalities}. We supply quantitative results and comparison of CSI with state-of-the art approaches on first a classical RL benchmark (the mountain car) Sec.~\ref{subsec:mountaincar} and then on a more difficult highway driving simulator (Sec.~\ref{subsec:highway}).
\subsection{Generalities}
\label{subsec:generalities}
\subsubsection{Data scarcity}
\label{subsubsec:data}
The CSI algorithm was designed not to repeatedly solve the RL problem. This feature makes it particularly well-suited to environments where sampling the MDP is difficult or costly. In the experiments, CSI is fed only with data sampled according to the expert policy. This corresponds for example to a situation where a costly system can only be controlled by a trained operator as a bad control sequence could lead to a system breakdown.

More precisely, the expert controls the system for $M$ runs of lengths $\{L_i\}_{1\leq i \leq M}$, %with values of $n$ in $\{3,7,10,15,20\}$,
giving samples $\{(s_k,a_k=\pi_E(a_k),s'_k)_k\} = D_E$. The dataset $D_C$ fed to the classifier is straightforwardly constructed from $D_E$ by dropping the $s'_k$ terms:
\begin{equation}
  D_C = \{(s_i = s_k,a_i=a_k)_i\}.
\end{equation}
\subsubsection{Heuristics}
\label{subsubsec:heuristics}

There is no particuliar problem for the classification step. It is however not reasonable to construct the dataset $D_R = \{((s_k,a_k=\pi_E(s_k)),\hat r_k)_k\}$ only from expert transitions and expect a small regression error term $\epsilon_R$. Indeed, the dataset $D_E$ only samples the dynamics induced by the expert's policy and not the whole dynamics of the MDP. 
This means that for a certain state $s_k$ we only know the corresponding expert action $a_k=\pi_E(s_k)$ and the following state $s'_k$ sampled according to the MDP dynamics : $s'_k \sim P(\cdot|s_k,a_k)$. For the regression to be meaningful, we need samples associating the same state $s_k$ and a different action $a\neq a_k$ with a datapoint $\hat r \neq \hat r_k$.\\
Recall that $\hat r_j = q(s_j,a_j) - \gamma q(s'_j,\pi_C(s'_j))$ (Eq.~\eqref{eq:rj}), without knowing $s'_k \sim P(\cdot|s_k,a\neq a_k)$, we can not provide the regressor with a datapoint to associate with $(s_k,a\neq a_k)$.
We will artificially augment the dataset $D_R$ with samples $((s_j=s_k,a),r_{min})_{j;\forall a\neq \pi_E(s_j) = a_k}$ where
\begin{equation}
  r_{min} = \min_k\hat r_k - 1.
\end{equation}
This heuristics intructs the regressor to associate a state-action tuple disagreeing with the expert ({\it i.e. }$(s_k,a\neq a_k)$) with a reward strictly inferior to any of those associated with expert state action tuples ({\it i.e. }$(s_k,a_k=\pi_E(s_k))$). Semantically, we are asserting that disagreeing with the expert in states the expert visited is a bad idea. This heuristics says nothing about states absent from the expert dataset. For such states the generalization capabilities of the regressor and, later on, the exploration of the MDP by an agent optimizing the reward will solve the problem.
Although this heuristics was not analyzed in Sec.~\ref{sec:analysis} (where the availability of a more complete dataset $D_R$ was assumed), the results shown in the next two subsections demonstrate its soundness.
\subsubsection{Comparison with state of the art approaches}
The similar looking yet fundamentally different algorithm SCIRL is an obvious choice as a contender to CSI as it advertises the same ability to work with very little data, without repeatedly solving the RL problem. In both experiments we give the exact same data to CSI and SCIRL.

@@

Finally, the classification policy $\pi_C$ output by the classification step of CSI was evaluated as well. Comparing classification and IRL algorithms makes no sense if the object of interest is the reward itself as can be envisionned in a biological or economical context. It is however perfectly sound to do so in an imitation context where what matters is the performance of the agent with respect to some objective criterion. Both experiments use such a criterion. Classification algorithms don't have to optimize any reward, the classification policy can directly be used in the environment. IRL algorithms on the other hand output a reward that must then be plugged in a RL or DP algorithm to get a policy. In each benchmark we used the same (benchmark-dependant) algorithm to get a policy from each of the three rewards output by SCIRL, CSI and RE. It is these policies whose performance we show. Finding a policy from a reward is of course a non-trivial problem that should not be swept under the rug, nevertheless we choose not to concern ourselves with it here as we wish to focus on IRL algorithms, not RL or DP algorithms. In this regard, using a classifier that directly outputs a policy may seem a much simpler solution, but we hope that the reader will be convinced that the gap in performance between classification and IRL is worth the trouble of solving the RL problem (once and for all, and not repeatedly as a subroutine like some other IRL algorithms).

@@

We referred in Sec.\ref{sec:algo} to \emph{score-function based multi-class classifiers} and explained how the classification rule is similar to the greedy mechanism that exists between an optimal action-value function and an optimal policy in an MDP. Most classification algorithm can be seen as such a classifier. In a simple $k$-nearest neighboor approach, for example, the score function $q(s,a)$ is the number of elements of class $a$ among the $k$-nearest neighboors of $s$. The generic M-SVM model makes the score function explicit (see \cite{guermeur2011generic}) (we use a SVM in the mountain car experiment Sec.~\ref{subsec:mountaincar}). In the highway experiment, we choose to use a structured margin classification approach \cite{taskar2005learning}.
%% that, given a feature vector $\phi: S\times A \rightarrow \mathbb{R}^p$ solves:
%% \begin{eqnarray*}
%%   &\min_{\theta,\zeta}\frac{1}{2}\|\theta\|^2 +
%%   \frac{\eta}{N}\sum_{i=1}^N \zeta_i \\
%%   &\text{~s.t.~} \forall i,
%%   \theta^T{\phi}(s_i,a_i)+\zeta_i \geq \max_a \theta^T
%%   {\phi}(s_i,a) + l(s_i,a). \label{eq:qp_taskar}
%% \end{eqnarray*}
%% with $l(s_i,a) = 1\textrm{ if } a\neq a_i \textrm{ and } l(s_i,a_i)=0$. The score function of this classifier is simply $q(s,a) = \theta^T \phi(s,a)$.
We chose a SVR as a regressor in the mountain car experiment and a simple least-square regressor on the highway.
%% Using the same features $\phi$ as our classifier, we introduce $\lambda$ a regularization coefficient (set to $0.1$ in our experiments), $I$ the identity matrix and 
%% \begin{eqnarray}
%%   \Phi = \begin{bmatrix}
%%     \vdots\\
%%     \phi(s_j,a_j)\\
%%     \vdots
%%   \end{bmatrix} \in \mathbb{R}^{n|A|\times p}\textrm{ and }
%%   \hat R = \begin{bmatrix}
%%     \vdots\\
%%     \hat r_j\\
%%     \vdots
%%   \end{bmatrix}\in \mathbb{R}^{n|A|},
%% \end{eqnarray}
%% with $r_j = r_{min}$ if $a_j \neq \pi_E(s_j)$. Our regressor can be written:
%% \begin{equation}
%%   \omega = (\Phi^T\Phi + \lambda I)^{-1}\Phi^T\hat R.
%% \end{equation}
%% The reward function is: $\hat R^C(s,a) = \omega^T \phi(s,a)$

@@

%% The feature function $\phi$ for this setting is the natural tabular feature function where each state action couple $(s,a)$ is associated with a unique index $i_{(s,a)}$. The only non-zero component of $\phi(s,a)$ is then the component at index $i_{(s,a)}$.

@@

Expert performance averages to $7.74$ ; we also show the natural random baseline that consists in drawing a random reward vector (with a uniform law) and training an agent on it. The reward functions found by SCIRL, CSI and RE are then optimized using a DP algorithm. The dataset $D_{random}$ needed by RE (and neither by CSI nor SCIRL) is made of $100$ episodes of length $10$ starting randomly in the state space and following a random policy. the dataset $D_E$ is made of $n$ episodes of length $n$, with $n \in \{3,7,10,15,20\}$.

Results are shown Fig.~\ref{fig:Highway} and Tab.~\ref{tab:highway}. We give the values of $\mathbf{E}_{s\sim\mathcal{U}}[V^{\pi}_{R_E}(s)]$ with $\pi$ being in turn the optimal policy for the rewards given by SCIRL, CSI and RE, the policy $\pi_C$ of the classifier (the very one the classification step of CSI outputs), and the optimal policy for a randomly drawn reward. Performance for CSI is slightly but definitively higher than for SCIRL (see the variance in Tab.~\ref{tab:highway}), slightly below the performance of the expert itself. Very few samples (100) are needed to reliably achieve expert-level performance.
